---
title: "Q1"
author: "Gowtham P" 
output: html_document
---


1.	Load the dataset “WA_Fn-UseC_-Marketing-Customer-Value-Analysis.csv" using pd.read_csv() and perform the following tasks with appropriate interpretation:

```{r}
library(dplyr)
library(ggplot2)
```

```{r}
#### 1. Load Data ####
df <- read.csv("D:/Data Science for Marketing-I& 2/dataset/WA_Fn-UseC_-Marketing-Customer-Value-Analysis.csv")
```
i. Perform basic exploratory data analysis (EDA) such as checking dataset shape and previewing the first few rows. What insights can be drawn from this initial exploration?

```{r}
head(df)
```
```{r}
dim(df)
```
```{r}
summary(df)
```
Interpertation:
The dataset has 9134 rows and 24 columns,head() function is used to display the first few rows of a dataset. 


ii. Analyze customer engagement by grouping data based on the Response variable. How does this grouping help in understanding customer behavior?

```{r}
# Encode engaged customers as 0s and 1s
df$Engaged <- rep(0,nrow(df))
df$Engaged[df$Response=='Yes']=1
```




```{r}
## - Overall Engagement Rates ##
engagementRate <- df %>% group_by(Response) %>%
  summarise(Count=n())  %>%
  mutate(EngagementRate=Count/nrow(df)*100.0)
```

iii. Visualize the engagement rate using a bar chart. What is the significance of this visualization, and how does the code achieve it?

```{r}

ggplot(engagementRate, aes(x=Response, y=EngagementRate)) +
  geom_bar(width=0.5, stat="identity") +
  ggtitle('Engagement Rate') +
  xlab("Engaged") +
  ylab("Percentage (%)") +
  theme(plot.title = element_text(hjust = 0.5)) 
```

Interpertation: 

Only 14.3% of customers responded positively, indicating a low engagement rate.
This suggests a need for improved marketing strategies to boost response rates.

iv. Calculate the engagement rate for different renewal offer types and interpret the results. Why is this metric useful?

```{r}
## - Engagement Rates by Offer Type ##
engagementRateByOfferType <- df %>% 
  group_by(Renew.Offer.Type) %>%
  summarise(Count=n(), NumEngaged=sum(Engaged))  %>%
  mutate(EngagementRate=NumEngaged/Count*100.0)
engagementRateByOfferType

ggplot(engagementRateByOfferType, aes(x=Renew.Offer.Type, y=EngagementRate)) +
  geom_bar(width=0.5, stat="identity") +
  ggtitle('Engagement Rates by Offer Type') +
  xlab("Offer Type") +
  ylab("Engagement Rate (%)") +
  theme(plot.title = element_text(hjust = 0.5)) 
```
Interpertation: 

Offer1 and Offer2 have the highest engagement rates (~16%), while Offer4 has the lowest (9.7%).
This suggests that some offers are more attractive, guiding future marketing strategies.

v. Extend the analysis by exploring engagement rates segmented by both Renew Offer Type and Vehicle Class. How does this multi-level grouping provide deeper insights?

```{r}
## - Offer Type & Vehicle Class ##
engagementRateByOfferTypeVehicleClass <- df %>% 
  group_by(Renew.Offer.Type, Vehicle.Class) %>%
  summarise(NumEngaged=sum(Engaged))  %>%
  left_join(engagementRateByOfferType[,c("Renew.Offer.Type", "Count")], by="Renew.Offer.Type") %>%
  mutate(EngagementRate=NumEngaged/Count*100.0)

ggplot(engagementRateByOfferTypeVehicleClass, aes(x=Renew.Offer.Type, y=EngagementRate, fill=Vehicle.Class)) +
  geom_bar(width=0.5, stat="identity", position = "dodge") +
  ggtitle('Engagement Rates by Offer Type & Vehicle Class') +
  xlab("Offer Type") +
  ylab("Engagement Rate (%)") +
  theme(plot.title = element_text(hjust = 0.5)) 
```
Interpertation:

More customers responded to Offer 2, especially those with Four-Door Cars. 
Offers 3 and 4 had very few responses.

vi. Perform customer segmentation using the variables ‘Customer Lifetime Value (CLV)’ and ‘Months Since Policy Inception’

```{r}

summary(df$Customer.Lifetime.Value)
summary(df$Months.Since.Policy.Inception)

clv_encode_fn <- function(x) {if(x > median(df$Customer.Lifetime.Value)) "High" else "Low"}
df$CLV.Segment <- sapply(df$Customer.Lifetime.Value, clv_encode_fn)

policy_age_encode_fn <- function(x) {if(x > median(df$Months.Since.Policy.Inception)) "High" else "Low"}
df$Policy.Age.Segment <- sapply(df$Months.Since.Policy.Inception, policy_age_encode_fn)

ggplot(
  df[which(df$CLV.Segment=="High" & df$Policy.Age.Segment=="High"),], 
  aes(x=Months.Since.Policy.Inception, y=log(Customer.Lifetime.Value))
) +
  geom_point(color='red') +
  geom_point(
    data=df[which(df$CLV.Segment=="High" & df$Policy.Age.Segment=="Low"),], 
    color='orange'
  ) +
  geom_point(
    data=df[which(df$CLV.Segment=="Low" & df$Policy.Age.Segment=="Low"),], 
    color='green'
  ) +
  geom_point(
    data=df[which(df$CLV.Segment=="Low" & df$Policy.Age.Segment=="High"),], 
    color='blue'
  ) +
  ggtitle('Segments by CLV and Policy Age') +
  xlab("Months Since Policy Inception") +
  ylab("CLV (in log scale)") +
  theme(plot.title = element_text(hjust = 0.5))
```
 Interpertation:
 
Customers are classified into High/Low CLV and Early/Late Policy Age groups.
This segmentation helps in prioritizing high-value customers for retention.


vii. Create a visualization to compare CLV against Months Since Policy Inception

```{r}
engagementRateBySegment <- df %>% 
  group_by(CLV.Segment, Policy.Age.Segment) %>%
  summarise(Count=n(), NumEngaged=sum(Engaged))  %>%
  mutate(EngagementRate=NumEngaged/Count*100.0)

ggplot(engagementRateBySegment, aes(x=CLV.Segment, y=EngagementRate, fill=Policy.Age.Segment)) +
  geom_bar(width=0.5, stat="identity", position = "dodge") +
  ggtitle('Engagement Rates by Customer Segments') +
  ylab("Engagement Rate (%)") +
  theme(plot.title = element_text(hjust = 0.5)) 
```

Customers with high CLV stay longer, so it's good to keep them happy.  
Customers with low CLV might leave early, so they need more attention.